Enforced Invariant Confluence

### 2018-09-19 doc/paper/sections/invariant_confluence.tex
Last tweaks!
Clarified read-only transactions.
From reviewer 11:

> It appears that there is an issue with readers who want to look at a
> dataset (e.g. compute an aggregate, or min). This triggers a merge?
> This issue should be clearly described.

Our formalism doesn't really discuss reads at all because reads don't
affect invariant confluence in any way. I'm guessing this is why the
reviewer got a little confused about what happens when you want to read
the object. I added a paragraph trying to clarify how invariant
confluence models reads.

A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of states
and a binary merge operator $\join: S \times S \to S$ which merges two states
into one. A \defword{transaction} $t: S \to S$ is a function which maps one
state to another. An \defword{invariant} $I$ is a subset of $S$.  Notationally,
we write $I(s)$ to denote that $s$ satisfies the invariant (i.e.  $s \in I$)
and $\lnot I(s)$ to denote that $s$ does not satisfy the invariant (i.e. $s
\notin I$).

\begin{example}\examplelabel{Ints}
  $O = (\ints, \max)$ is a distributed object consisting of integers merged by
  $\max$; $t(x) = x + 1$ is a transaction which adds one to a state; and
  $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are
  non-negative.
\end{example}

@@

We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
valid execution of our system model. Formally, we define
$\sTIreachablepredicate{e}$ to be the smallest predicate that satisfies the
following equations:
\begin{itemize}
  \item
    $\sTIreachablepredicate{s_0}$.
  \item
    For all expressions $e$ and for all transactions $t$ in the set $T$ of
    transactions, if $\sTIreachablepredicate{e}$ and $I(t(e))$, then
    $\sTIreachablepredicate{t(e)}$.
  \item
    For all expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
    $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
    e_2}$.
\end{itemize}
Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists a
\sTIreachable{} expression $e$ that evaluates to $s$. Returning to
\exampleref{Ints} with start state $s_0 = 42$, we see that all integers greater
than or equal to 42 (i.e.\ $\setst{x \in \ints}{x \geq 42}$) are
\sTIreachable{}, and all other integers are \sTIunreachable{}.

>>>>

A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of states
and a binary merge operator $\join: S \times S \to S$ that merges two states
into one. A \defword{transaction} $t: S \to S$ is a function that maps one
state to another. An \defword{invariant} $I$ is a subset of $S$. Notationally,
we write $I(s)$ to denote that $s$ satisfies the invariant (i.e. $s \in I$)
and $\lnot I(s)$ to denote that $s$ does not satisfy the invariant (i.e. $s
\notin I$).

\begin{example}\examplelabel{Ints}
  $O = (\ints, \max)$ is a distributed object consisting of integers merged by the
  $\max$ function; $t(x) = x + 1$ is a transaction that adds one to a state; and
  $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are
  non-negative.
\end{example}

Note that by modelling a transaction $t$ as a function $S \to S$, we focus
exclusively on the effects that a transaction has on the object (i.e.\
``writes'' to the object). Transactions are also free to read the value of the
object, but these reads are not captured by our model because, as we'll see,
they do not affect \invariantconfluence{}. For example, we could model any
read-only transaction as a function $t$ where $t(s) = s$ for every $s \in S$.

@@

We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
valid execution of our system model. Formally, we define
$\sTIreachablepredicate{e}$ to be the smallest predicate that satisfies the
following equations:
\begin{itemize}
  \item
    $\sTIreachablepredicate{s_0}$.
  \item
    For all expressions $e$ and for all transactions $t$ in the set $T$ of
    transactions, if $\sTIreachablepredicate{e}$ and $I(t(e))$, then
    $\sTIreachablepredicate{t(e)}$.
  \item
    For all expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
    $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
    e_2}$.
\end{itemize}
Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists an
\sTIreachable{} expression $e$ that evaluates to $s$. Returning to
\exampleref{Ints} with start state $s_0 = 42$, we see that all integers greater
than or equal to 42 (i.e.\ $\setst{x \in \ints}{x \geq 42}$) are
\sTIreachable{}, and all other integers are \sTIunreachable{}.
